---
title: "PAD_data_cleaning_code"
format: html
editor: visual
---

```{r}
# Library required packages
library(tidyverse)
library(readxl)
```

# Unredactions

The following section cross-references usaspending, IATI, and fa.gov data to reduce the redaction rate in foreign assistance data

## IATI Data

Loads in IATI data for 2016 - 2022 for Colombia, Guatemala and Mexico

## FA.gov Data

Load in fa.gov dataset and get a list of redacted implementers

```{r}
# Load data
fa_data <- read_csv("./data/fa.gov_8-15-2024.csv") %>%
  filter(
    `Fiscal Year` %in% c(2016:2022),
    `Managing Agency ID` == 1,
    `Aid Type Group Name` == "Project-Type"
  )

# Get a list of redacted implementers

# List of implementer tags that signify redactions
redacted_imps <- c(
  "NGO - International Redacted", 
  "Enterprise - Non United States Redacted",
  "Enterprise - Non United States Other",
  "NGO - Non United States Redacted",
  "Multilateral Organizations - Redacted",
  "Church and Faith Based - Non United States Redacted", 
  "Unspecified Vendors - Colombia",
  "Unspecified Vendors - Mexico", 
  "Unspecified Vendors - Peru",
  "NGO - Non United States Other",
  "Unspecified Vendors - El Salvador",
  "Church and Faith Based - United States Redacted",
  "Enterprise - United States Redacted",
  "Multilateral - United Nations Organization Redacted",
  "NGO - United States Redacted",
  "University/Research Institute - United States Redacted", 
  "Multilateral - United Nations Organization Redacted"
)

# Get list of redacted projects fa.gov
redacted_df <- fa_data %>%
  filter(`Implementing Partner Name` %in% redacted_imps) 
```

## USASpending

Cross-references usaspending data with fa.gov

```{r}
# Load usaspending contracts and assistance for USAID 
contracts <- read_csv("./data/Contracts_PrimeAwardSummaries_2024-08-18_H21M40S21_1.csv")
assistance <- read.csv("./data/Assistance_PrimeAwardSummaries_2024-08-18_H21M40S38_1.csv")
contracts_id <- contracts %>% select(award_id_piid, recipient_name, recipient_parent_name, recipient_country_code)
assistance_id <- assistance %>% select(award_id_fain, recipient_name, recipient_parent_name, recipient_country_code) 
# Rename id column to bind data
names(assistance_id)[names(assistance_id) == 'award_id_fain'] <- 'award_id_piid'
project_ids <- rbind(contracts_id, assistance_id)
```

Join USASpending and redacted fa.gov data

```{r}
redacted_projs_final <- redacted_df %>%
  # Standardize ID's for join 
  mutate(`Activity Project Number` = str_remove_all(
    `Activity Project Number`,
    "-"
  )) %>%
  select(`Country Code`, `Implementing Partner Category Name`, `Implementing Partner Sub-category ID`, `Activity Name`, `Activity Description`, `Activity Project Number`) %>%
left_join(project_ids, 
          by = join_by("Activity Project Number" == "award_id_piid"))
```

This dataframe of redacted projects (redacted_projs_final) is further cleaned manually since some projects are not correctly tagged in USASpending as USAID projects

```{r}
# Load dataset with additional manual cleaning
usaspending_unredactions <- read_excel("./data/usaspending_unredactions.xlsx") %>%
  select(`Activity Project Number`, recipient_name, recipient_country_code) %>%
  group_by(`Activity Project Number`) %>%
  summarize(
    recipient_name = max(recipient_name),
    recipient_country_code = max(recipient_country_code)
  ) %>%
  filter(`recipient_name` != "NA")
```

```{r}
# Join with fa.gov data
unredacted_fa_data <- fa_data %>%
  # Standardize IDs for join
  mutate(`Activity Project Number` = str_remove_all(
      `Activity Project Number`,
      "-"
  )) %>%
  left_join(
    usaspending_unredactions,
    by = join_by(`Activity Project Number`)
  ) %>% 
  mutate(
    `Implementing Partner Name` = case_when(
      !is.na(recipient_name) ~ recipient_name,
      .default =  `Implementing Partner Name`
    )
  ) 
```

Check redaction rate

```{r}
# Base redaction rate
fa_data %>%
  filter(`Implementing Partner Name` %in% redacted_imps) -> test

sum(test$`Current Dollar Amount`) / sum(fa_data$`Current Dollar Amount`)

# Redaction rate after cross referencing with USASpending
unredacted_fa_data %>%
  filter(`Implementing Partner Name` %in% redacted_imps) -> test

sum(test$`Current Dollar Amount`) / sum(fa_data$`Current Dollar Amount`)
```

## IATI Data

```{r}
# Load in IATI DATA
iati_df <- read_excel("./data/iati_9_11_co_guat_mex.xlsx") %>%
  # Make years tidy
  pivot_longer(cols = c(10:16), names_to = "year") %>%
  # Filter awards with value = 0
  filter(value != 0) %>%
  # Clean years column
  mutate(year = str_replace(year, "Value \\(USD\\)", "")) %>%
  mutate(year = str_replace_all(year, "\\(", "")) %>%
  mutate(year = as.numeric(str_replace_all(year, "\\)", ""))) %>%
  # Filter to projects
  filter(`Aid Type` %in% c("C01 - Project-type interventions", "A02 - Sector budget support"))
```

```{r}
# Clean fa.gov data

duplicates = c(
  "BANCO DE LAS MICROFINANZAS BANCAMIA S A" = "Bancama",
  "LA CORPORACION MISION DE OBSERVACION ELECTORAL MOE" = "Misin de Observacin Electoral (Colombia)", 
  "SECRETARIADO NACIONAL DE PASTORAL SOCIAL-CARITAS COLOMBIANA" = "Caritas Colombiana", 
  "FUNDACION PARA LA IMPLEMENTACION, DISENO, EVALUACION Y ANALISIS DE POLITICAS PUBLICAS, A.C." = "Fundacion IDEA",
  "MEXICO EVALUA. CENTRO DE ANALISIS DE POLITICAS PUBLICAS, A.C." = "Mexico Evalua Centro de Analisis de Politicas Publicas A.C.",
  "ABT GLOBAL LLC" = "Abt Associates, Inc.",
  "ACDI/VOCA" = "Agriculture Cooperative Development International/Volunteers in Overseas Cooperative Assistance",
  "AMERICARES FOUNDATION, INC." = "Americares",
  "ARCANGELES FUNDACION PARA LA REHABILITACION INTEGRAL" = "Arcangeles",
  "BANYAN GLOBAL INC." = "Banyan Global",
  "COLLEGE OF WILLIAM AND MARY" = "College of William & Mary",
  "CHEMONICS INTERNATIONAL, INC." = "Chemonics International, Inc.",
  "DAI GLOBAL LLC" = "Development Alternatives, Inc.",
  "DEXIS INTERACTIVE LLC" = "Dexis Consulting Group",
  "ENVIRONMENTAL INCENTIVES, LLC" = "Environmental Incentives, LLC",
  "GLOBAL COMMUNITIES, INC." = "Global Communities",
  "HEARTLAND ALLIANCE INTERNATIONAL, LLC" = "Heartland Alliance",
  "JHPIEGO CORP" = "Jhpiego Corporation",
  "MERCY CORPS" = "Mercy Corps",
  "MIYAMOTO INTERNATIONAL INC" = "Miyamoto Relief",
  "PAN AMERICAN DEVELOPMENT FOUNDATION, INC." = "Pan-American Development Foundation",
  "PRONATURA SUR, A.C." = "Pronatura Mexico", 
  "SAVE THE CHILDREN FEDERATION, INC." = "Save the Children Federation, Inc.", 
  "SOCIAL IMPACT INC" = "Social Impact, Inc.", 
  "SSG-ADVISORS LLC" = "SSG Advisors, LLC", 
  "UNIVERSITY OF NORTH CAROLINA AT CHAPEL HILL" = "University of North Carolina, Chapel Hill",
  "WORLD VISION INC." = "World Vision"
)

unredacted_fa_data <- unredacted_fa_data %>%
  mutate(`Implementing Partner Name` = ifelse(
    `Implementing Partner Name` %in% names(duplicates),
    duplicates[`Implementing Partner Name`], # IF duplicate
    `Implementing Partner Name` # IF not
  ))
```

```{r}
# Clean IATI IDS
iati_df %>%
  mutate(
    `IATI Identifier` = str_replace(`IATI Identifier`, "US-GOV-1-", "")
    ) %>%
  mutate(
    `IATI Identifier` = str_replace_all(`IATI Identifier`, "-", "")
  ) -> iati_ids_cleaned

# Group FA data for join
grouped_fa_data <- unredacted_fa_data %>%
  group_by(`Activity Project Number`) %>%
  summarize(
    imp_name = max(`Implementing Partner Name`)
  )

# join the df
fa_iati_joined <-left_join(iati_ids_cleaned, grouped_fa_data, by = join_by("IATI Identifier" == "Activity Project Number"))
```

```{r}
# check redaction rate

# redacted imps list 
redacted_list_iati <- c(
  "Unspecified Vendors - Colombia", 
  "USAID redacted this field in accordance with the exceptions outlined in the Foreign Aid Transparency and Accountability Act of 2016.",
  "USAID redacted this field in accordance with the exceptions outlined in the Foreign Aid Transparency and Accountability Act of 2016. [nan]",
  "Unspecified Vendors - Guatemala"
)

# IATI Redaction Rate
iati_df %>%
  filter(`Receiver Organisation` %in% redacted_list_iati) -> iati_redacted
sum(iati_redacted$value) / sum(iati_df$value)
# The reaction rate of IATI data is 0.03108498

# joined redaction rate
fa_iati_joined %>%
  filter(`Receiver Organisation` %in% redacted_list_iati) %>%
  filter(is.na(imp_name)) -> redacted
sum(redacted$value) / sum(fa_iati_joined$value)
# The reaction rate of IATI data joined with fa.gov is 0.00315654
```

## Join Cleaned Data With Coding

```{r}
org_coding <- read_xlsx("./data/org_codings.xlsx") %>%
  select(-Note)

# Org Type Coding - Generates sheet: Org. Type Coding COL-GUATE-MX
org_type_coding_df <- fa_iati_joined %>%
  mutate(redacted_indicator = case_when(
    `Receiver Organisation` %in% redacted_list_iati & !is.na(`imp_name`) ~ "T",
    .default = "F"
  )) %>%
  mutate(`Receiver Organisation` = case_when(
    `Receiver Organisation` %in% redacted_list_iati & !is.na(`imp_name`) ~ `imp_name`,
    .default = `Receiver Organisation`
  )) %>% ## appends unredactions if available and corrects names if needed
  left_join(org_coding, by = join_by(`Receiver Organisation` == `imp_name`)) %>%
  mutate(`Receiver Organisation` = case_when(
     !is.na(`Correct_name`) ~ `Correct_name`,
     TRUE ~ `Receiver Organisation`
  )) %>%
  select(-imp_name, -`Correct_name`) 

# Direct Imp Coding Df
direct_imp_coding <- org_type_coding_df %>%
  select(-Code, -Organization_Type, -redacted_indicator)
```
